---
title: "Analysis of Power Generators in the US"
output: html_notebook
author: "Ryan Willett"
---

Loading libraries
```{r, echo=F}
library(shiny)
library(shinydashboard)
library(tidyverse)
library(ggplot2)
library(plotly)
library(leaflet)
library(googleVis)

# Initialization of plot theme
t <- theme_classic()
```

Loading data
```{r}
data <- read.csv("./data/january_generator2019.csv", stringsAsFactors = F, header = T)
data <- data %>% rename(
  capacity_MW = Nameplate.Capacity..MW.,
  capacity_MW_summer = Net.Summer.Capacity..MW.,
  capacity_MW_winter = Net.Winter.Capacity..MW.) %>% 
  select(-Google.Map,-Bing.Map) %>% 
  mutate(capacity_MW = gsub(",","", capacity_MW), 
         capacity_MW_summer = gsub(",","", capacity_MW_summer),
         capacity_MW_winter = gsub(",","", capacity_MW_winter))

colsToNum <- c("capacity_MW", "capacity_MW_summer", "capacity_MW_winter")
data[colsToNum] <- sapply(data[colsToNum], as.numeric)
```

## Tab 1

Map enabling filtering of various features including: 
* Location of all generators

```{r}

#Testing first with all nuclear
data_nuc <- data %>% filter(Technology == "Natural Gas Steam Turbine")

plant_map <- leaflet(data = data_nuc) %>% addTiles() %>%
  addMarkers(clusterOptions = markerClusterOptions() ,
             lng = ~Longitude,
             lat = ~Latitude,
             popup = ~Plant.Name)
plant_map

```

```{r}
AllStates <- map("state", fill = TRUE,
                 plot = FALSE,
                 region = c(
                   "Alabama",
                   "Alaska",
                   "American Samoa",
                   "Arizona",
                   "Arkansas",
                   "California",
                   "Colorado",
                   "Connecticut",
                   "Delaware",
                   "District of Columbia",
                   "Florida",
                   "Georgia",
                   "Hawaii",
                   "Idaho",
                   "Illinois",
                   "Indiana",
                   "Iowa",
                   "Kansas",
                   "Kentucky",
                   "Louisiana",
                   "Maine",
                   "Maryland",
                   "Massachusetts",
                   "Michigan",
                   "Minnesota",
                   "Mississippi",
                   "Missouri",
                   "Montana",
                   "Nebraska",
                   "Nevada",
                   "New Hampshire",
                   "New Jersey",
                   "New Mexico",
                   "New York",
                   "North Carolina",
                   "North Dakota",
                   "Ohio",
                   "Oklahoma",
                   "Oregon",
                   "Pennsylvania",
                   "Rhode Island",
                   "South Carolina",
                   "South Dakota",
                   "Tennessee",
                   "Texas",
                   "Utah",
                   "Vermont",
                   "Virginia",
                   "Washington",
                   "West Virginia",
                   "Wisconsin",
                   "Wyoming"
                 ))
```

```{r}
# data %>% group_by(Plant.State) %>% summarise(sum_MW = sum(capacity_MW)) %>% arrange(Plant.State)
# 
# leaflet() %>% addTiles() %>% addPolygons(data = AllStates,
#                                          #fillColor = heat.colors(8, alpha = 1),
#                                          color = "black",
#                                          stroke = FALSE)
```

* Quantify that technology by state

Google Vis chart that visualizes that amount of power in MW that each state produces with that power technology
```{r}
tech_byState <- data %>% filter(Technology == "Nuclear") %>% group_by(Plant.State) %>% summarise(sum_MW = sum(capacity_MW)) %>% as.data.frame()
#tech_byState <- tech_byState[-1,]

techbystate <- gvisGeoChart(tech_byState[-1,], "Plant.State", "sum_MW",
                 options=list(region="US", displayMode="regions",
                              resolution="provinces",
                              width="auto", height="auto"))
plot(techbystate)
```

#### Tab 2 ####

* What is the production of the power by technology in the selected state.

```{r}
data %>% 
```


* Showing generators of particular type

```{r}

# Breakdown of the number of generators total by technology
data %>% group_by(Technology) %>% tally() %>% arrange(desc(n))

data %>% group_by(Technology) %>% summarise(sum_MW = sum(capacity_MW)) %>% arrange(desc(sum_MW))
```

* Aggregated summary of total number of generators in each state
```{r}
num_per_state <- data %>% group_by(Plant.State) %>% tally()

ggplot(num_per_state, aes(x = reorder(Plant.State, desc(Plant.State)), y = n)) + t +
  geom_col() + coord_flip() +
  labs(x = "State", y = "Number of Power Generators")
```

Bar graph displaying the breakdown of:
* Number of each generator type in the state
* Amount of power produced  by each power technology in the state 
* The company with the most number of generators
* The companies ranked in the state by power production
```{r}
data %>% 
  filter(Plant.State == "CA") %>% group_by(Technology) %>% tally() %>% arrange(desc(n)) %>% 
  ggplot(aes(reorder(Technology, desc(Technology)), y = n)) + t +
  geom_col() + coord_flip() +
  labs(x = "Technology", y = "Number of Power Generators")

data %>% filter(Plant.State == "CA") %>% group_by(Technology) %>% summarise(sum_MW = sum(capacity_MW)) %>% arrange(desc(sum_MW)) %>% 
  ggplot(aes(reorder(Technology, desc(Technology)), y = sum_MW)) + t +
  geom_col() + coord_flip() +
  labs(x = "State", y = "Power Production (MW)")

data %>% filter(Plant.State == "CA") %>% group_by(Entity.Name) %>% tally() %>% filter(n==max(n))

data %>% filter(Plant.State == "CA") %>% group_by(Entity.Name) %>% summarise(sum_MW = sum(capacity_MW)) %>% filter(sum_MW == max(sum_MW))
```

* Number of power generator by company
```{r}
data %>% group_by(Technology) %>% 
```

```{r}
ggplot(data, aes(x=Technology, y=Operating.Year, col=Technology)) + 
  geom_jitter() +
  theme_classic()
```

